Kung Faux Pandas \\ \large Simplifying privacy protection

### 2018-06-04 paper/main.tex
some edits

% this tool simplfies data sharing, simplifies data exploration within organizations
\begin{abstract}
A description of an end-to-end system for easily generating faux-data which is statistically similar to given real data but does not contain sensitive personal information. This system, which is publicly available \footnote{\url{https://github.com/CUD2V/kungfauxpandas}}, enables data to be distributed for reproducibility testing and other purposes while in full compliance with HIPAA and GDPR.
\end{abstract}

@@

The ability of independent investigators to verify results is perhaps \emph{the} most fundamental feature that separates science from other forms intellectual inquiry. Unfortunately, the mechanisms of communicating scientific results have changed little since the age of enlightenment and are inadequate for enabling reproducibility in some modern scientific endeavors such as machine learning. Perhaps the most important impediment to reproducibility is the need to share and have access to data. This has long been difficult in health care fields, but will presently become difficult for just about everyone.

There is a large body of published literature detailing methods for overcoming this difficulty related to data via de-identification (removing key identifiers, grouping individuals, and related anonymization techniques \cite{hippapro}) and synthesizing data (creating data through algorithmic means, population level statistics, randomization, and ideally preserving inter-relationships between records \cite{walonoski_synthea_2018, patki_synthetic_2016, choi_generating_2017}), however at present a mechanism to routinely \emph{use} these techniques has not been forthcoming.

This paper details an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular combination of published privacy protection methods with the popular Python Pandas data science library\cite{mckinney-proc-scipy-2010}. KFP also provides a Structured Query Language (SQL) interface which enables users to arbitrarily query any data within a database, returning to the user a data set which either does not contain any personal information though synthetic means or anonymization. We collectively refer to de-identified data and synthetic data as \emph{faux data}. Faux in it's standard sense of artificial or fake, but more specifically as it is used in the fashion sense of the word: the use of man made materials that look and feel like real leather, fur, etc., but wihout the problematic derivation from live animals and reliance upon limited resources.

@@

% shorten this? Is this the right place in the background for this?
 Over the past 18 years, there has been several significant papers clarifying the definition and importance of reproducibility along with the closely related term replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result" \cite{leek_opinion_2015}; others such as Drummond use the terms in a reverse fashion \cite{drummond_replicability_2009}. Despite the semantic disagreement, both groups agree that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing details key factors in reproducibility as having: data, input parameters, documentation, software code, and an environment capable of running the provided software code; items that cannot easily nor clearly be communicated by the traditional research paper \cite{schwab_making_2000}. The previously mentioned authors propose that reproducibility should be treated as a minimum required standard that all published research should meet \cite{peng_reproducible_2006}. Specifically in the context of machine learning, this means including details to a level someone else could create the same environment including hardware configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

@@

% What is our major contribution to the area of reproducibility in machine learning?
Kung Faux Pandas sidesteps this problem by generating \emph{faux-data} that is ``statistically similar'' to the real data but does not itself contain any sensitive data.

>>>>

% this tool simplfies data sharing, simplifies data exploration within organizations
\begin{abstract}
Kung Faux Pandas is an end-to-end system for easily generating synthetic data that is statistically similar to real data but lacks sensitive information. This publicly available \footnote{\url{https://github.com/CUD2V/kungfauxpandas}} system lowers barriers to HIPAA- and GDPR-compliant data sharing for reproducibility testing and other purposes.
\end{abstract}

@@

Independent reproduction and replication of results are critical components of scientific inquiry. Barriers to data access and sharing are the most important impediments to reproducibility in computational research. Health care research has unique challenges because of necessary personal data privacy protections.

Several methods exist to de-identify (remove key identifiers, group individuals, and related anonymization techniques \cite{hippapro}) and synthesize data (create data through algorithmic means, population level statistics, randomization, ideally with preservation of multivariable relationships between records \cite{walonoski_synthea_2018, patki_synthetic_2016, choi_generating_2017}). However, no pipeline currently exists with which computational investigators can routinely \emph{anonymize} or \emph{synthesize} data to facilitate access and sharing. This work addresses that gap.

We have created an open source software library called \emph{Kung Faux Pandas (KFP)} which allows for the modular combination of established privacy protection methods with the popular Python Pandas data science library\cite{mckinney-proc-scipy-2010}. KFP also provides a Structured Query Language (SQL) interface which enables users to query a database and receive a data set without personal information through either anonymization or synthetic data generation.

@@

% shorten this? Is this the right place in the background for this?
Recent work has clarified the definition and importance of reproducibility and a closely related entity, replicability. Leek and Peng have defined reproducibility as the "ability to recompute data analytic results given an observed dataset and knowledge of the data analysis pipeline," and replicability as "the chance that an independent experiment targeting the same scientific question will produce a consistent result."\cite{leek_opinion_2015,peng_reproducible_2006} Others have used the terms in a reverse fashion.\cite{drummond_replicability_2009} Despite the semantic differences, there is broad agreement that an independent experiment with confirmatory results is the strongest support of any experiment. One early influential paper on the topic of reproducibility in scientific computing lists the key factors in reproducibility as available data, input parameters, documentation, software code, and an environment capable of running the provided software code. These items are difficult to completely convey in a traditional research paper \cite{schwab_making_2000}. 
In the context of machine learning (ML), sufficient detail is required such that an independent investigator could create the same hardware and software configuration, data used for all experiments, source code, documentation on data and how to run/configure software, and tests that verify the software runs correctly. Once a result can be reproduced, new researchers can then build upon the methods, gather new data for testing/validation, or discover alternative methods to replicate a result.

@@

% What is our major contribution to the area of reproducibility in machine learning?
KFP sidesteps this problem by generating \emph{faux-data} that is ``statistically similar'' to the real data but does not itself contain any sensitive data. We collectively refer to de-identified data and synthetic data as \emph{faux data}. Faux in it's standard sense of artificial or fake, but more specifically as it is used in the fashion sense of the word: the use of man made materials that look and feel like real leather, fur, etc., but wihout the problematic derivation from live animals and reliance upon limited resources.
